Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2018-09-25 SupportingInformation.Rmd
modified

output: 
  pdf_document:
    number_sections: yes
    keep_tex: yes
    fig_caption: yes
    includes:
      in_header: head_supp.sty

@@

\tableofcontents
\listoftables
\listoffigures

@@

```{r read_BMR_data, echo=FALSE, results="markup"}
set.seed(seed)
#---------------------------------------------------------
BMR_dat<-read.csv("./Data/BMR/bmr-contrasts.csv",header = T)
#class(BMR_dat)
#knitr::kable(head(BMR_dat[,c(1,4,5)] ,3),format='pandoc',
#             caption = "First few rows of raw BMR data.\\label{tab_rawdata_BMR}",
#             booktabs=T) # display first few rows
```

```{r plot_BMR_data, echo=FALSE, results="hide"}
set.seed(seed)
source("phylo_dat_fn.R")

# Generating raw data plot btw ln(mass) and BMR from bmr-contrasts.csv
pdf("./Results/BMR_results/rawdataplot_lnBMR_vs_lnmass.pdf")
op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
plot(BMR_dat$ln.mass,BMR_dat$ln.BMR,type="p",col="red",xlab="ln(mass) of contrasts",ylab="ln(BMR) of contrasts",
     cex.lab=1.5,cex.axis=1.5)
par(op)
dev.off()
#--------------------------------------------------------------------------------------------
# Generating copula plot btw ln(mass) and ln(BMR) from bmr-contrasts.csv
pdf("./Results/BMR_results/copula_lnBMR_vs_lnmass.pdf")
v_BMR<-phylo_dat_fn(data=BMR_dat,i=4,j=3,ploton = T) 
#plot(v_BMR[,1],v_BMR[,2],type="p",col="red",xlab="ln(mass)",ylab="BMR",
#     cex.lab=1.5,cex.axis=1.5)
dev.off()
```

```{r read_phylo_data, echo=F, results="markup"}
set.seed(seed)
phylo_dat<-read.csv("./Data/Anolis/contrasts.csv",header = T)
#class(phylo_dat) #it's a data.frame : no NA/missing value in this dataset
#tab_phylo<-phylo_dat[,c(1,7,8)] 
#knitr::kable(head(tab_phylo,3),format='pandoc',
#             caption = "First few rows of raw phylo data.\\label{tab_rawdata_phylo}",
#             booktabs=T) # display first few rows
```

```{r plot_phylo_data, echo=F, results="hide"}
set.seed(seed)
source("phylo_dat_fn.R")
#----------------------------------------------------------------------------------------------------
# Generating all copula plots between different pathlength and traits 
pdf("./Results/Anolis_results/contrasts_csv_results/contrasts_csv_pathlength_traits_cop_plot.pdf",width=20, height = 20)
op<-par(mfrow=c(4,4))
i<-1
  for(j in c(1:13)){
    v_phylo<-phylo_dat_fn(data=phylo_dat,i=i,j=j,ploton = T)
    mtext(paste0(" [",i,",",j,"] ", ","," n=",dim(v_phylo)[1]),side = 3,col="blue")
  } 
par(op)
dev.off()
#------------------------------------------------------------------------------------------------------------
# Generating all 12 by 12 copula plots between different traits
pdf("./Results/Anolis_results/contrasts_csv_results/contrasts_csv_btw_traits_cop_plot.pdf",width=20, height = 20)
op<-par(mfrow=c(12,12))
for(i in c(2:13)){
  for(j in c(2:13)){
    v_phylo<-phylo_dat_fn(data=phylo_dat,i=i,j=j,ploton = T)
    mtext(paste0(" [",i,",",j,"] ", ","," n=",dim(v_phylo)[1]),side = 4,col="blue",line=0.01)
  } 
}
par(op)
dev.off()
#----------------------------------------------------------------------------------------
# Generating raw data plot btw AVG.ltoe.IV and AVG.humerus from contrasts.csv
pdf("./Results/Anolis_results/contrasts_csv_results/rawdataplot_AVG_ltoe_humerus.pdf")
op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
plot(phylo_dat$AVG.ltoe.IV,phylo_dat$AVG.humerus,type="p",col="red",xlab="contrasts for AVG.ltoe.IV",ylab="contrasts for AVG.humerus",
     cex.lab=1.5,cex.axis=1.5)
par(op)
dev.off()
#--------------------------------------------------------------------------------------------
# Generating copula plot btw AVG.ltoe.IV and AVG.humerus from contrasts.csv
pdf("./Results/Anolis_results/contrasts_csv_results/copula_AVG_ltoe_humerus.pdf")
v_contrasts_67<-phylo_dat_fn(data=phylo_dat,i=6,j=7,ploton = T)
dev.off()
```

@@

<!--# Nonparametric analysis of tail dependence \label{NPA}

Left (or lower)- and right (or upper)-tail dependence quantify the extent to which two ranked
variables $u_i$, $v_i$ for $i=1,...,n$ 
(see \nameref{MT-Model_selection} for defintions of $u_i$ and $v_i$ in the manuscript)
are related to each other in the left and right tails, respectively, of their 
distributions [@nelsen2007_copula, @joe2014_dependence]. 
Our nonparametric study of tail dependence uses three
statistics which can quantify the extent to which $u_i$ and $v_i$ are 
related in any part of their distributions. We here describe the 
statistics. The statistics were formulated with positively associated
variables in mind. All the variables we study are positively associated
when they are significantly associated. So definitions and associated 
diagrams will assume positive associations, though generalizations are 
possible. Given two bounds 
$0 \leq l_b < u_b \leq 1$, we define the parallel lines $u+v=2l_b$ and 
$u+v=2u_b$ which intersect the unit square in which the $u_i$ and $v_i$ 
fall (these are the blue and red lines, respectively, in 
Fig. \ref{fig_stat_image}). Our statistics quantify the dependence
between $u_i$ and $v_i$ in the region bounded by these lines.
Using $l_b=0$ and some value $u_b\leq 0.5$ can give information about 
dependence in the left parts of the distributions of $u$ and $v$, and 
using $u_b=1$ and $l_b \geq 0.5$ can give information about dependence in 
the right parts of the distributions.

\begin{figure}[!h]
\begin{center}
\includegraphics[width=6 cm,height=6 cm]{./stat_image.jpg} \hspace{0.6 cm}
\caption{Schematic diagram supporting defintions of nonparametric tail dependence statistics \label{fig_stat_image}}
\end{center}
\end{figure}

The first quantity, $\cor_{l_b,u_b}(u,v)$, is the portion of the Spearman 
correlation of $u_i$ and $v_i$ that is attributable to the points 
lying in the region given by the bounds $u+v=2l_b$ and $u+v=2u_b$, i.e.,
\begin{equation}\label{eq.Cor}
   \cor_{l_b,u_b}(u,v) = \frac{1}{(n-1)\sqrt{\var(u)\var(v)}}\sum 
   (u_i-\mean(u)) (v_i-\mean(v)),
\end{equation}
where means and variances are computed using all $n$ data 
points, but the
sum is over only the indices $i$ for which $u_i+v_i > 2l_b$ and 
$u_i+v_i < 2u_b$. 
Larger values of $\cor_{l_b,u_b}$ indicate stronger positive association
between $u$ and $v$ in the region given by the bounds.

We also defined a statistic $\Ps_{l_b,u_b}$, using the same bounds
$u+v=2l_b$ and $u+v=2u_b$, as follows. For a distance $h$, we defined
$S(h)$ to be the number of points $(u_i,v_i)$ within the bounds 
and a distance less than $h$ from the line $v=u$, divided by the 
total number of points within the bounds. This function is
defined for $h$ going from $0$ to a distance $h_{\text{max}}$ which
is half the longer of the two segments obtained by intersecting the
bounds with the unit square. It is easy to see $S(0)=0$ and 
$S(h_{\text{max}})=1$. We defined $S_i(h)$, for the same range 
of $h$, to be the area within the bounds and the unit square
and within distance $h$ of the line $v=u$, divided by the area
within the bounds and the unit square. This is the expected value of
$S(h)$ for independent data. We then defined
\begin{equation}\label{eq.P}
   P_{l_b,u_b} = \int_0^{h_{max}} ( S(h) - S_i(h) ) dh.
\end{equation}
As for $\cor_{l_b,u_b}$, larger values of $\Ps_{l_b,u_b}$ 
indicate stronger positive association
between $u$ and $v$ in the region given by the bounds.

The third statistic, $\Dsq_{l_b,u_b}(u,v)$, is the average squared 
distance between points satisfying $u_i+v_i>2l_b$ and 
$u_i+v_i<2u_b$ and the line $v=u$. Unlike $\cor_{l_b,u_b}$ and 
$\Ps_{l_b,u_b}$, which are large when points satisfying $u_i+v_i>2l_b$ and 
$u_i+v_i<2u_b$ cluster close to the line $v=u$, $\Dsq_{l_b,u_b}$ is 
then small. 
Thus large values of $\cor_{l_b,u_b}$ and $\Ps_{l_b,u_b}$ indicate strong 
dependence in the portions of the distributions of $u$ and $v$ given by 
the bounds $u_i+v_i>2l_b$ and $u_i+v_i<2u_b$, whereas small values of 
$\Dsq_{l_b,u_b}$ indicate strong dependence.

For large datasets (i.e., datasets with large $n$),
we can consider $l_b$ and $u_b$ close together without incurring undue
sampling variation in our statistics, and we can consider 
several different bands $(l_b,u_b)$ to understand how dependence
varies in different parts of the distributions. But for datasets with 
smaller $n$ we considered only $l_b=0$, $u_b=0.5$ and $l_b=0.5$, $u_b=1$.
We use the shorthand $\cor_l=\cor_{0,0.5}$ ($l$ is for "lower") 
and $\cor_u=\cor_{0.5,1}$ ($u$ is for "upper"); likewise $\Ps_l=\Ps_{0,.5}$,
$\Ps_u=\Ps_{0.5,1}$, $\Dsq_l=\Dsq_{0,0.5}$, and $\Dsq_u=\Dsq_{0.5,1}$.

The difference $\cor_l-\cor_u$ is expected to be positive if 
dependence in the left halves of the distributions is stronger 
than dependence in their right halves; likewise for $\Ps_l-\Ps_u$.
This holds, for instance, for the Clayton copula [@nelsen2007_copula].
The difference $\Dsq_u-\Dsq_l$, in which the "upper" and
"lower" statistics occur in opposite order, is also expected to be 
positive under the same conditions. We use these three differences
to detect asymmetry of tail dependence. Likewise, for large datasets,
we consider $\cor_{l_b,u_b}-\cor_{1-u_b,1-l_b}$, 
$\Ps_{l_b,u_b}-\Ps_{1-u_b,1-l_b}$ and
$\Dsq_{1-u_b,1-l_b}-\Dsq_{l_b,u_b}$ with $l_b=0$ and $u_b$ close to $0$.

We tested our statistics by testing whether $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ revealed the asymmetry of tail dependence known to be present in data simulated from various copulas (for details see Appendix \ref{Test_npa_stat}).

For bivariate datasets, we compared values 
of the statistics $\cor_{l_b,u_b}$,
$\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, and $\Dsq_u-\Dsq_l$ to 
distributions of values of the same statistics computed on surrogate datasets that were produced from 
the empirical data by randomizing it in a special way. The surrogate/randomized datasets had exactly 
the same marginal distributions as the empirical data and had very similar Kendall or Spearman 
correlation (the algorithm had two versions, one for preserving each correlation coefficient), 
but had Normal copula structure. Thus our comparisons tested the null hypothesis 
that our statistics took values on the empirical data no different from what would have been expected 
if the copula structure of the data were Normal, but the data were otherwise 
statistically unchanged.
Details of how surrogates were prroduced are in Appendix \ref{surrog_test}, and code is in [`copsurrog2d`](https://github.com/sghosh89/BIVAN/blob/master/copsurrog2d.R).

The comparison of one of the
statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, $\Dsq_{l_b,u_b}$, as computed on the 
empirical data, to the distribution
of its values computed on many surrogate datasets 
provides a test of whether dependence in the part of the distribution specified 
by $l_b$ and $u_b$ is different from
what would be expected from a null hypothesis of normal copula structure but 
unchanged marginals and unchanged 
Kendall (or Spearman)
correlation. Thus the comparison helps answer the first two parts of question \hyperref[MT-Intro]{\textbf{(Q1)}} 
from the \nameref{MT-Intro} for bivariate datasets. 
Significant deviations correspond to deviations from normal 
copula structure. In particular, deviations 
using $l_b=0$ and $u_b$ small (say, $0.1$) correspond to lower tail dependence 
different from that of a normal copula;
likewise, using $l_b=0.9$ and $u_b=1$ tests for upper tail dependence different 
from that of a normal copula.

The comparison of one of the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, $\Dsq_u-\Dsq_l$, 
as computed on the 
empirical data, to the distribution
of its values on many surrogate datasets 
provides a test of asymmetry of tail dependence. Thus this comparison helps answer the
third part of question \hyperref[MT-Intro]{\textbf{(Q1)}} from the \nameref{MT-Intro} for bivariate datasets.

For multivariate datasets, a spatial resampling scheme was used instead of the surrogate 
approach described above.
The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
widely used spline correlogram method; the same scheme was also used by @walteretal2017 
as part of a method for 
performing model selection among matrix regression models. For each pair of 
locations $i \neq j$, the statistic
of interest, here denoted $s_{ij}$ ($\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, 
$\Dsq_{l_b,u_b}$, $\cor_l-\cor_u$, 
$\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$) was computed for those locations. The 
values $s_{ij}$ were then arranged in
a matrix, $S$, with diagonal entries NA. The mean of the non-NA entries of $S$ was computed. For 
$\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$, or $\Dsq_{l_b,u_b}$ this mean characterized the 
average strength of 
dependence, across all pairs of distinct locations, for data in the portion of the distributions 
given by $l_b$ and $u_b$. For $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, or $\Dsq_u-\Dsq_l$, the mean characterized average
asymmetry of tail dependence across all pairs of locations. 
Confidence intervals for the values of $\mean(S)$ were then computed as quantiles of a 
distribution of bootstrapped values. If $\mean(S)$ was significantly different from 0, as
revealed by these confidence intervals, for the statistics $\cor_l-\cor_u$, $\Ps_l-\Ps_u$, 
or $\Dsq_u-\Dsq_l$, then tail dependence was significantly asymmetric, in contrast to a 
Normal null hypothesis, and answering question \hyperref[MT-Intro]{\textbf{(Q1)}} from the \nameref{MT-Intro}.
Bootstrapping was performed by resampling the locations of measurement, with replacement. 
This corresponds to 
resampling, with replacement, the rows and columns of $S$, and is described in detail elsewhere 
[@bjornstadfalck2001 ; @walteretal2017].-->

@@

<!--
# Question 1 analyses for each dataset
Recall that question 1 is: Do datasets in ecology and related fields have non-Gausian copula structure? Do they show tail dependence distinct from that of a Normal copula? In particular do they show asymmetric tail dependence?

## Details of methods \label{Method}
### Model selection approach \label{Model_selection}
### Non-parametric analysis of tail dependence \label{NPA}
moved to Paper.Rmd
-->

<!--
## Results: soil C and N data \label{Ex_SoilCN}
### Model selection approach \label{Ex_SoilCN_Model_selection} 
### Non-parametric analysis of tail dependence \label{Ex_SoilCN_NPA}
                    [[[[moved to Paper.Rmd]]]]
-->

@@

<!--
\newpage
## Results: BMR data \label{Ex_BMR}
### Model selection approach \label{Ex_BMR_Model_selection}
### Non-parametric analysis of tail dependence \label{Ex_BMR_NPA}
                  ----moved to Paper.Rmd----
-->

<!--This chunk is for model selection results of BMR data-->
```{r bivMS_BMR, echo=F, cache=T, cache.extra=list(seed,v_BMR,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
set.seed(seed) 
source("OurBiCopSelect.R")
#families<-c(1,3:10,13:14,16:20) 
BivMS_BMR<-OurBiCopSelect(u1=v_BMR[,1],u2=v_BMR[,2],families=c(1,3:10,13:14,16:20),
                          level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,gofnormal=F,status=F)
saveRDS(BivMS_BMR,file="./Results/BMR_results/BivMS_lnBMR_vs_lnmass.RDS")
```

<!--Display model selection results for BMR data in a table-->
```{r tab_BMRfit, echo=F, results='markup'}
h<-BivMS_BMR$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
knitr::kable(h,
             format='pandoc',
             caption = "Fitting info for **BMR dataset**. \\label{tab_BMRfit}",
             booktabs=T)
```

<!--This chunk is for npa testing results of BMR dataset-->
```{r stat_BMR_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,v_BMR,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
source("bivfunctionplot.R")
set.seed(seed) 
# Now do NPA for v_BMR
stat_BMR_ga<-bivfunctionplot(v=v_BMR,resloc="./Results/BMR_results/",nametag="BMR_numbin_10",numbin = 10)
stat_BMR_ga_numbin5<-bivfunctionplot(v=v_BMR,resloc="./Results/BMR_results/",nametag="BMR_numbin_5",numbin = 5)
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/BMR_results/BMR_numbin_10_bivfunctionplot.pdf}
\caption[Testing non-parametric statistics for BMR data with $numbin=10$]{Nonparametric tests for tail dependence and other deviations from a normal copula for \textbf{BMR data}. As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red cross), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (blue and green cross show 0.025 and 0.975 quantiles of these results).  When red cross was outside blue and green cross, text at the top of plots indicates how many surrogate values the data value was less than or greater than. \label{fig_BMR_nonparam}}
\end{center}
\end{figure}

<!--
\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/BMR_results/BMR_numbin_5_bivfunctionplot.pdf}
\caption[Testing non-parametric statistics for BMR data with $numbin=5$]{Nonparametric tests for tail dependence and other deviations from a normal copula for BMR data. As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.2$, $0.2-0.4$, ..., $0.8-1$ (red lines), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (dashed lines show 0.025 and 0.975 quantiles of these results and dotted lines show 0.005 and 0.995 quantiles). \label{fig_BMR_nonparam5}}
\end{center}
\end{figure}
-->

```{r tab_BMR_asym, echo=F, results="markup"}
tab_BMR_asym<-make_table_stat_fn(data=stat_BMR_ga)
tab_BMR_asym_numbin5<-make_table_stat_fn(data=stat_BMR_ga_numbin5)

knitr::kable(tab_BMR_asym,
             format="pandoc",
             caption = "**BMR data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis. \\label{tab_BMR_asym}",
             booktabs=T)

#knitr::kable(tab_BMR_asym_numbin5,
#             format="pandoc",
#             caption = "Results for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis (best fit is here Frank with symmetric tail dep)?, numbin=5 \\label{tab_BMR_asym_numbin5}",
#             booktabs=T)
```

<!--
\newpage
## Results: contrasts.csv data \label{Ex_contrasts}
### Model selection approach \label{Ex_contrasts_Model_selection}
### Non-parametric analysis of tail dependence \label{Ex_contrasts_67_NPA}
                 ----moved to Paper.Rmd----
-->

<!--This chunk is for model selection results of anolis data-->
```{r bivMS_contrasts_67, echo=F, cache=T, cache.extra=list(seed,v_contrasts_67,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
set.seed(seed)
# Now do Model Selection for [i,j]=[6,7]
# 6 : AVG. ltoe.IV
# 7 : AVG. humerus
source("OurBiCopSelect.R")
families<-c(1,3:10,13:14,16:20)
BivMS_contrasts_67<-OurBiCopSelect(u1=v_contrasts_67[,1],u2=v_contrasts_67[,2],families=families,
                          level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
                          gofnormal=F,status=F)
saveRDS(BivMS_contrasts_67,file="./Results/Anolis_results/contrasts_csv_results/BivMS_contrasts_67.RDS")
```

<!--Display model selection results for contrasts data in a table-->
```{r tab_contrasts_67fit, echo=F, results='markup'}
h<-BivMS_contrasts_67$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
knitr::kable(h,
             format='pandoc',
             caption = "Fitting info for **Anolis data set** with contrasts for AVG.ltoe.IV and AVG.humerus phylogenetic traits. \\label{tab_contrasts_67fit}",
             booktabs=T)
```

<!--This chunk is for npa testing results of anolis dataset-->
```{r stat_contrasts_67_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,v_contrasts_67,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}   
source("bivfunctionplot.R")  
set.seed(seed) 
# Now do NPA for v_contrasts_67 with numbin=4
stat_contrasts_67_ga<-bivfunctionplot(v=v_contrasts_67,resloc="./Results/Anolis_results/contrasts_csv_results/",
                                      nametag="contrasts_67",numbin = 4) 
```

```{r stat_contrasts_67_graphic_approach2, echo=F, results="hide", cache=T, cache.extra=list(seed,v_contrasts_67,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
source("bivfunctionplot.R")
set.seed(seed) 
# Now do NPA for v_contrasts_67

stat_contrasts_67_ga_numbin2<-bivfunctionplot(v=v_contrasts_67,
                            resloc="./Results/Anolis_results/contrasts_csv_results/",
                            nametag="contrasts_67_numbin2",numbin = 2)

stat_contrasts_67_ga_numbin3<-bivfunctionplot(v=v_contrasts_67,
                            resloc="./Results/Anolis_results/contrasts_csv_results/",
                            nametag="contrasts_67_numbin3",numbin = 3)

stat_contrasts_67_ga_numbin5<-bivfunctionplot(v=v_contrasts_67,
                            resloc="./Results/Anolis_results/contrasts_csv_results/",
                            nametag="contrasts_67_numbin5",numbin = 5)

stat_contrasts_67_ga_numbin6<-bivfunctionplot(v=v_contrasts_67,
                            resloc="./Results/Anolis_results/contrasts_csv_results/",
                            nametag="contrasts_67_numbin6",numbin = 6)

stat_contrasts_67_ga_numbin7<-bivfunctionplot(v=v_contrasts_67,
                            resloc="./Results/Anolis_results/contrasts_csv_results/",
                            nametag="contrasts_67_numbin7",numbin = 7)

stat_contrasts_67_ga_numbin10<-bivfunctionplot(v=v_contrasts_67,
                            resloc="./Results/Anolis_results/contrasts_csv_results/",
                            nametag="contrasts_67_numbin10",numbin = 10)
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/Anolis_results/contrasts_csv_results/contrasts_67_bivfunctionplot.pdf}
\caption[Testing non-parametric statistics for anolis data]{Nonparametric tests for tail dependence and other deviations from a normal copula for anolis data. As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.25$, $0.25-0.5$, $0.5-0.75$, $0.75-1$ for $4$ number of bins (red cross), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (blue and green cross show 0.025 and 0.975 quantiles of these results). \label{fig_contrasts_67_nonparam}}
\end{center}
\end{figure}

```{r tab_contrasts_67_asym, echo=F, results="markup"}
tab_contrasts_67_asym<-make_table_stat_fn(data=stat_contrasts_67_ga)
knitr::kable(tab_contrasts_67_asym,
             format="pandoc",
             caption = "**Anolis data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis \\label{tab_contrasts_67_asym}",
             booktabs=T)
```

<!--
This chunks are just to see the tables for all numbin variation for anolis data set: It appears that numbin=4 is optimized, when numbin=2, >= 7, there is no significant asymetry in tail dep in all 3 types of stats.

```{r tab_contrasts_67_asym67_numbin2, echo=F, results="markup"}
#tab_contrasts_67_asym2<-make_table_stat_fn(data=stat_contrasts_67_ga_numbin2)
#knitr::kable(tab_contrasts_67_asym2,
#            format="pandoc",
#             caption = "numbin_2,anolis \\label{tab_contrasts_67_asym2}",
#             booktabs=T)
```

```{r tab_contrasts_67_asym67_numbin3, echo=F, results="markup"}
#tab_contrasts_67_asym3<-make_table_stat_fn(data=stat_contrasts_67_ga_numbin3)
#knitr::kable(tab_contrasts_67_asym3,
#             format="pandoc",
#             caption = "numbin_3,anolis \\label{tab_contrasts_67_asym3}",
#             booktabs=T)
```

```{r tab_contrasts_67_asym67_numbin5, echo=F, results="markup"}
#tab_contrasts_67_asym5<-make_table_stat_fn(data=stat_contrasts_67_ga_numbin5)
#knitr::kable(tab_contrasts_67_asym5,
#            format="pandoc",
#             caption = "numbin_5,anolis \\label{tab_contrasts_67_asym5}",
#             booktabs=T)
```

```{r tab_contrasts_67_asym67_numbin6, echo=F, results="markup"}
#tab_contrasts_67_asym6<-make_table_stat_fn(data=stat_contrasts_67_ga_numbin6)
#knitr::kable(tab_contrasts_67_asym6,
#             format="pandoc",
#             caption = "numbin_6,anolis \\label{tab_contrasts_67_asym6}",
#             booktabs=T)
```

```{r tab_contrasts_67_asym67_numbin7, echo=F, results="markup"}
#tab_contrasts_67_asym7<-make_table_stat_fn(data=stat_contrasts_67_ga_numbin7)
#knitr::kable(tab_contrasts_67_asym7,
#             format="pandoc",
#             caption = "numbin_7,anolis \\label{tab_contrasts_67_asym7}",
#             booktabs=T)
```

```{r tab_contrasts_67_asym67_numbin10, echo=F, results="markup"}
#tab_contrasts_67_asym10<-make_table_stat_fn(data=stat_contrasts_67_ga_numbin10)
#knitr::kable(tab_contrasts_67_asym10,
#             format="pandoc",
#             caption = "numbin_10,anolis \\label{tab_contrasts_67_asym10}",
#             booktabs=T)
```

-->

@@

\newpage
<!--Display model selection results for Ceder-creek data for year 2000 in a table-->
```{r tab_cc2000fit, echo=F, results='markup'}
h<-BivMS_res_cc$`2000`$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
knitr::kable(h,
             format='pandoc',
             caption = "Fitting info for **Ceder-creek dataset** for year 2000. \\label{tab_cc2000fit}",
             booktabs=T)
```

@@

<!--
\newpage
## Results: abundance data for the green spruce aphid, *Elatobium abietinum* \label{Ex_count}
### Model selection approach \label{Ex_count_Model_selection}
### Non-parametric analysis of tail dependence \label{Ex_count_NPA}
                  ----moved to Paper.Rmd----
-->

@@

<!--
\newpage
## Results: first flight date data for the leaf-curling plum aphid, *Brachycaudus helichrysi* \label{Ex_ff}
### Model selection approach \label{Ex_ff_Model_selection}
### Non-parametric analysis of tail dependence  \label{Ex_ff_NPA}
                          ----moved to Paper.Rmd----
-->

@@

<!--
\newpage
## Results: methane flux data \label{Ex_methane}
### Model selection approach \label{Ex_methane_Model_selection}
### Non-parametric analysis of tail dependence  \label{Ex_methane_NPA}
                 ----moved to Paper.Rmd----
-->

@@

  m<-matrix(NA,1,12)
  colnames(m)<-c("Indep.p","AIC-best.copula","Highest.AICw",
                 "N.AICw","p.CvM","p.KS",
                 "LT_model.avg","UT_model.avg","(LT-UT)_model.avg",
                 "Rank_Pl_K","Rank_Pu_K","Rank_Pl-Pu_K")
  m<-as.data.frame(m)

>>>>

output: 
  pdf_document:
    number_sections: yes
    keep_tex: yes
    fig_caption: yes
    toc: True
    includes:
      in_header: head_supp.sty

@@

<!--\tableofcontents-->
\listoftables
\listoffigures

@@

```{r read_BMR_birds_and_copulaplot, echo=F, results='hide'}
set.seed(seed)
raw_b<-readxl::read_xls("./Data/BMR/BirdBodyMassesMetabolicRates/BirdMassesAndMetabolicRatesMcNab2008.xls")
raw_b<-raw_b[,1:3]
raw_b<-na.omit(raw_b)
raw_b<-raw_b[,2:3]
raw_b<-as.data.frame(raw_b)
raw_b[303,2]<-0.66  #to convert as numeric
raw_b[,2]<-as.numeric(raw_b[,2])
raw_b<-raw_b[-which(raw_b[,1]<0),]  # omit -ve mass?

b<-log10(raw_b)
colnames(b)<-c("log10M(g)","log10BMR(KJ/h)")
pdf("./Results/BMR_results/birds/raw_birds_BMR.pdf")
op<-par(mar=c(5.1, 4.1, 0.2, 2.1))
plot(b[,1],b[,2],type="p",col="red",xlab="log(Mass)",ylab="log(BMR)",cex.lab=1.5,cex.axis=1.5)
par(op)
dev.off()

source("getcopula.R")
pdf("./Results/BMR_results/birds/copula_birds_BMR.pdf")
cop_b<-getcopula(d=b,rankon=T,ploton=T) #convert to a copula and plot
dev.off()
```

@@

@@

@@

```{r bivmodselect_birds_BMR, echo=F, cache=T, cache.extra=list(seed,cop_b,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"),mtime("preprocessors.R"))}
source("OurBiCopSelect.R")
set.seed(seed)
BivMS_birds<-OurBiCopSelect(u1=cop_b[,1],u2=cop_b[,2],families=c(1,3:10,13:14,16:20),
                              level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,
                              gofnormal=F,status=F)
saveRDS(BivMS_birds,file="./Results/BMR_results/birds/BivMS_birds.RDS")
```

<!--Display model selection results for BMR:birds data in a table-->
```{r tab_bmr_birds_fit, echo=F, results='markup'}
h<-BivMS_birds$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
knitr::kable(h,
             format='pandoc',
             caption = "Fitting info for **birds' BMR dataset** . \\label{tab_bmr_birds_fit}",
             booktabs=T)
```

```{r stat_bmr_birds_graphic_approach, echo=F, results="hide", cache=T, cache.extra=list(seed,cop_b,mtime("bivfunctionplot.R"),mtime("CopulaFunctions_flexible.R"),mtime("copsurrog2d.R"))}
source("bivfunctionplot.R")
set.seed(seed)
stat_birds_ga<-bivfunctionplot(v=cop_b,resloc="./Results/BMR_results/birds/",nametag="birds_numbin_10",numbin = 10)
```

\begin{figure}[!h]
\begin{center}
\includegraphics[width=16 cm]{./Results/BMR_results/birds/birds_numbin_10_bivfunctionplot.pdf}
\caption[Testing non-parametric statistics for Ceder-creek data with $numbin=10$ for year 2000]{Nonparametric tests for tail dependence and other deviations from a normal copula for \textbf{birds' BMR data}. As described in \nameref{MT-Methods}, each statistic was computed on the data using the ranges $0-0.1$, $0.1-0.2$, ..., $0.9-1$ (red cross), and was also computed using the same ranges for 1000 normal-copula surrogate datasets (blue and green cross show 0.025 and 0.975 quantiles of these results).  When red cross was outside blue and green cross, text at the top of plots indicates how many surrogate values the data value was less than or greater than. \label{fig_bmr_birds_nonparam}}
\end{center}
\end{figure}

```{r tab_bmr_birds_asym, echo=F, results="markup"}
tab_bmr_birds_asym<-make_table_stat_fn(data=stat_birds_ga)
knitr::kable(tab_bmr_birds_asym,
             format="pandoc",
             caption = "**Birds' BMR data set results** for tests of whether asymmetry of tail dependence was significant compared to a normal-copula null hypothesis on year 2000 \\label{tab_bmr_birds_asym}",
             booktabs=T)
```

@@

<!--Display model selection results for Ceder-creek data for year 2000 in a table-->
```{r tab_cc2000fit, echo=F, results='markup'}
h<-BivMS_res_cc$`2000`$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
knitr::kable(h,
             format='pandoc',
             caption = "Fitting info for **Ceder-creek dataset** for year 2000. \\label{tab_cc2000fit}",
             booktabs=T)
```

@@

@@

@@

@@

  m<-matrix(NA,1,12)
  colnames(m)<-c("Indep.p","AIC-best.copula","Highest.AICw",
                 "N.AICw","p.CvM","p.KS",
                 "LT.model.avg","UT.model.avg","(LT-UT).model.avg",
                 "Rank.Pl.K","Rank.Pu.K","Rank.Pl-Pu.K")
  m<-as.data.frame(m)

  m$Indep.p<-round(r$IndepTestRes,3)
  m$`AIC-best.copula`<-r$InfCritRes$copname[which.min(r$InfCritRes$AIC)]
  m$Highest.AICw<-round(max(r$InfCritRes$AICw),3)
  m$N.AICw<-round(r$InfCritRes$AICw[1],3)
  m$p.CvM<-r$GofRes_CvM
  m$p.KS<-r$GofRes_KS
  m$LT.model.avg<-round(r$relLTdep_AICw,3)
  m$UT.model.avg<-round(r$relUTdep_AICw,3)
  m$`(LT-UT).model.avg`<-m$LT.model.avg-m$UT.model.avg
  m$Rank.Pl.K<-rs$Rank_Pl_K
  m$Rank.Pu.K<-rs$Rank_Pu_K
  m$`Rank.Pl-Pu.K`<-rs$Rank_PlmPu_K
  mt<-as.data.frame(t(m))
  return(mt)
